upwell_IDX

### 2020-04-20 upwell_IDX.Rmd
Some comments

In this equation μ represents the wind speed (m/s) and θ represents the wind direction which is measured in degrees. The 160 degrees is used as this refers to the angle of the coastline. This equation is largely dependant on wind speed and direction data in order to determing the intensity of the upwelling event. Wind data were obtained daily from the South African Weather Service (SAWS). This wind data was then matched to the date at which temperature were collected. With this data the upwelling index was determined.

@@

```{r}
# Installling andloading libraries

@@

```{r}
# wind_1 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings = "",
#                      col.names = c("station_number", "station_name", "date", "hour", "sub", "speed", "dir"))
# 
# wind_2 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings = "",
#                      col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
# 
# wind_3 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings = "",
#                      col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
# 
# # Slecting the important columns for each of the datasets
# wind_fix <- function(df){
# wind <- df %>%
#   select(station_name, date, hour, dir, speed) %>%  #column names
#   mutate(date = as.Date(as.character(date)),
#          hour = as.numeric(as.character(hour)),
#          dir = as.numeric(as.character(dir)),
#          speed = as.numeric(as.character(speed)))
# }
# 
# wind_1 <- wind_fix(df = wind_1)
# wind_2 <- wind_fix(df = wind_2)
# wind_3 <- wind_fix(df = wind_3)
# 
# ## Renaming the sites within the wind datasets to match the name of the sites at which seawater temperature was collected
# ## The wind data was obtained from the SAWS and the wind stations used were the closes stations to which temperature was collected
# 
# renaming_sites_1 <- function(df) {
#   sites <- df %>%
#       mutate(temp_sites = case_when(station_name == "CAPE TOWN TABLE BAY" ~ "Sea Point",
#                                     station_name == "CAPE TOWN - ROYAL YACHT CLUB" ~ "Sea Point",
#                                     station_name  == "PORT NOLLOTH" ~"Port Nolloth",
#                                     station_name  == "CAPE TOWN SLANGKOP" ~ "Hout Bay",
#                                     station_name  == "LAMBERTSBAAI NORTIER" ~ "Lamberts Bay",
#                                     station_name  == "LANGEBAANWEG AWS" ~ "Saldanha Bay"))
#   return(sites)
# }
# 
# 
# wind_sitesmatched_1 <-  renaming_sites_1(df = wind_1)
# wind_sitesmatched_2 <-  renaming_sites_1(df = wind_2)
# wind_sitesmatched_3 <-  renaming_sites_1(df = wind_3)
# wind_data <- rbind(wind_sitesmatched_3,wind_sitesmatched_2,wind_sitesmatched_1)
# save(wind_data, file = "Data/wind_data.RData")
# 
# load("Data/wind_data.RData")
# selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay", "Hout Bay")
# 
# wind_daily <- wind_data %>%
#   # ungroup() %>%
#   # mutate(date = as.Date(date))%>% # Changing the date to date format
#   dplyr::group_by(temp_sites, date) %>%
#   filter(temp_sites %in% selected_sites) %>%  # Matching the sites to wind temp sites
#   dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees")),2), # Direction using the circular function
#                    mean_speed = round(mean(speed),2)) # Calculating the mean
# save(wind_daily, file = "Data/wind_daily.RData")
```

@@

```{r}
# Coast R- function that calculates the angle of the site perpendicular to the coastline
# Transect function in coastR package finds the angle away or along the coastline
south_africa_away_wide <- coastR::transects(site_list_sub, spread = 30)

@@

# Create titles
titles <- c("Shore-normal", "Islands")
# Plotting function
plot_sites <- function(site_list, buffer, title_choice, dist){

@@

  # Visualise
  world_map +
    geom_segment(data = site_list, colour = "red4", 
                 aes(x = lon, y = lat, xend = lon_dest, yend = lat_dest)) +
    geom_point(data = site_list, size = 3, colour = "black", aes(x = lon, y = lat)) +
    geom_point(data = site_list, size = 3, colour = "red", aes(x = lon_dest, y = lat_dest)) +
    coord_cartesian(xlim = c(min(site_list$lon - buffer), 
                             max(site_list$lon + buffer)),
                    ylim = c(min(site_list$lat - buffer), 
                             max(site_list$lat + buffer))) +
    labs(x = "", y = "", colour = "Site\norder") +
    ggtitle(titles[title_choice])
}
south_africa_away <- plot_sites(south_africa_away_wide, 1, 1, 100000)
south_africa_away
```

@@

```{r}
UI_angle <- wind_daily %>%  # Making refeerence to the wind daily datta created above
  dplyr::rename(site = sites) %>%  # Renaming sites to site
  left_join(south_africa_away_wide[,c(2,21)], by = "site") %>%  # Joining wind data to south africa wide data column 2 and 21. by site 
  dplyr::rename(coast_angle = heading) %>% # Renamingthe column heading to coast_angle
  mutate(ui.saws = mean_speed * (cos(dir_circ - coast_angle))) %>% # applying the ui formula
  drop_na %>% # removing the na values
  dplyr::rename(t = date) #renaming date to t

@@

# Exceedence function heatwaveR - This is done to determine the consecutive number of days at or above what the UI value is meant to be   # Detect consecutive days in exceedance of a given threshold.

@@

# Bits of code testing the thresholds against the upwelling literature
##################### Upwelling ######################
# Min duration for an upwelling event - 1day and threshold -1. Upwelling can over within a few hours, but given that we have daily data - upwelling has a minimum duration of 1 day
# Seasonal upwelling. Upwelling seasons are summer and spring
# Did the wind intensify? This is usually done by taking first third of the dataset and the last thid of the dataset and doing a t-test on this
# Strong south easterly winds lead to a stronger upwelling event
##############################################

>>>>

In this equation μ represents the wind speed (m/s) and θ represents the wind direction which is measured in degrees. The 160 degrees is used as this refers to the angle of the coastline. This equation is largely dependant on wind speed and direction data in order to determining the intensity of the upwelling event. Wind data were obtained daily from the South African Weather Service (SAWS). This wind data were then matched to the date at which temperatures were collected. With these data the upwelling index was determined.

@@

```{r}
# Installing and loading libraries

@@

```{r, eval=FALSE}
wind_1 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings = "",
                     col.names = c("station_number", "station_name", "date", "hour", "sub", "speed", "dir"))

wind_2 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings = "",
                     col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))

wind_3 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings = "",
                     col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))

# Selecting the important columns for each of the datasets
## Renaming the sites within the wind datasets to match the name of the sites at which seawater temperature was collected
## The wind data was obtained from the SAWS and the wind stations used were the closes stations to which temperature was collected
wind_fix <- function(df){
wind <- df %>%
  mutate(date = as.Date(as.character(date)),
         hour = as.numeric(as.character(hour)),
         dir = as.numeric(as.character(dir)),
         speed = as.numeric(as.character(speed)),
         temp_sites = case_when(station_name == "CAPE TOWN TABLE BAY" ~ "Sea Point",
                                station_name == "CAPE TOWN - ROYAL YACHT CLUB" ~ "Sea Point",
                                station_name  == "PORT NOLLOTH" ~"Port Nolloth",
                                station_name  == "CAPE TOWN SLANGKOP" ~ "Hout Bay",
                                station_name  == "LAMBERTSBAAI NORTIER" ~ "Lamberts Bay",
                                station_name  == "LANGEBAANWEG AWS" ~ "Saldanha Bay")) %>% 
  select(temp_sites, date, hour, dir, speed) %>% 
  drop_na()
}

wind_fix_1 <- wind_fix(df = wind_1)
wind_fix_2 <- wind_fix(df = wind_2)
wind_fix_3 <- wind_fix(df = wind_3)

# Combine and save
wind_data <- rbind(wind_fix_3, wind_fix_2, wind_fix_1)
save(wind_data, file = "Data/wind_data.RData")

load("Data/wind_data.RData")
selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay", "Hout Bay")

# RWS: Please figure out why some wind direction values are returned as NA
# RWS: Also, it's not enough to just calculate mean speed. It needs to be taken into account with the directions.
# RWS: Please look through these resources and make corrections accordingly:
# https://cran.r-project.org/web/packages/rWind/rWind.pdf
# https://www.kaggle.com/srlightfoote/example-wind-analysis-in-r
wind_daily <- wind_data %>%
  # ungroup() %>%
  # mutate(date = as.Date(date))%>% # Changing the date to date format
  filter(temp_sites %in% selected_sites) %>%  # Matching the sites to wind temp sites
  dplyr::group_by(temp_sites, date) %>%
  dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees", 
                                                           template = 'geographics')),2), # Direction using the circular function
                   mean_speed = round(mean(speed),2)) # Calculating the mean
save(wind_daily, file = "Data/wind_daily.RData") 
```

@@

```{r}
# Coast R- function that calculates the angle of the site perpendicular to the coastline
# Transect function in coastR package finds the angle away or along the coastline
south_africa_away_wide <- coastR::transects(site_list_sub, spread = 30) %>% 
  mutate(heading = ifelse(heading < 0, heading+360, heading))

@@

# Plotting function
plot_sites <- function(site_list, buffer, dist){

@@

  # Visualise
  world_map +
    geom_segment(data = site_list, colour = "red4", 
                 aes(x = lon, y = lat, xend = lon_dest, yend = lat_dest)) +
    geom_point(data = site_list, size = 3, colour = "black", aes(x = lon, y = lat)) +
    geom_point(data = site_list, size = 3, colour = "red", aes(x = lon_dest, y = lat_dest)) +
    coord_cartesian(xlim = c(min(site_list$lon - buffer), 
                             max(site_list$lon + buffer)),
                    ylim = c(min(site_list$lat - buffer), 
                             max(site_list$lat + buffer))) +
    labs(x = "", y = "", colour = "Site\norder")
}
south_africa_away <- plot_sites(south_africa_away_wide, 1, 100000)
south_africa_away
```

@@

```{r}
UI_angle <- wind_daily %>%  # Making reference to the wind daily data created above
  dplyr::rename(site = sites) %>%  # Renaming sites to site
  left_join(south_africa_away_wide[,c(2,21)], by = "site") %>%  # Joining wind data to South Africa wide data column 2 and 21. by site 
  drop_na() %>% # removing the na values
  dplyr::rename(coast_angle = heading) %>% # Renaming the column heading to coast_angle
  mutate(ui.saws = mean_speed * (cos(dir_circ - coast_angle))) %>% # applying the ui formula
  dplyr::rename(t = date) #renaming date to t

@@

# Exceedence function heatwaveR 
This is done to determine the consecutive number of days at or above what the UI value is meant to be   # Detect consecutive days in exceedance of a given threshold.

@@
